### 2023-11-24 chapters/paper_retrospective.tex
Add citation to top reference
Update initial reference
Update top reference
Convert retrospective from EN-UK to EN-US

\textit{This chapter is a piece of original research published as part of the project:} \newline
\begin{center}
    \begin{enumerate}[leftmargin=8mm,rightmargin=8mm,topsep=0mm,label={[\Alph*]}]
        \setcounter{enumi}{5}
        \item \fullcite{wenstrup_retrospective_2023} \shared
    \end{enumerate}
\end{center}

@@

\section*{Abstract}
Advanced stroke treatment is time dependent and, therefore, relies on recognition by call-takers at prehospital telehealth services to ensure fast hospitalisation. This study aims to develop and assess the potential of machine learning in improving prehospital stroke recognition during medical helpline calls. We use calls from 1 January 2015 to 31 December 2020 in Copenhagen to develop a machine learning-based classification pipeline. Calls from 2021 are used for testing. Calls are first transcribed using an automatic speech recognition model and then categorized as stroke or non-stroke using a text classification model. Call-takers achieve a sensitivity of 52.7\% (95\% confidence interval 49.2-56.4\%) with a positive predictive value (PPV) of 17.1\% (15.5-18.6\%). The machine learning framework performs significantly better (p < 0.0001) with a sensitivity of 63.0\% (62.0-64.1\%) and a PPV of 24.9\% (24.3-25.5\%). Thus, a machine learning framework for recognising stroke in prehospital medical helpline calls may become a supportive tool for call-takers, aiding in early and accurate stroke recognition.

\section{Introduction}
Stroke is a leading cause of disability and death worldwide \parencite{cite1,cite2,cite3}. Effective treatment is time-sensitive, and an optimal outcome is more likely when treatment is administered within the first four and a half hours from stroke onset \parencite{cite4,cite5}. The gateway to ambulance transport and hospital admittance is through prehospital telehealth services, including emergency medical call centres, nurse advice call lines, and out-of-hours health services. In the pre-hospital setting, the use of mobile stroke units has made it possible to deliver advanced treatment faster \parencite{cite6,cite7}. As the mobile stroke unit is only dispatched to patients with a suspected stroke, the impact of mobile stroke unit is directly influenced by accurate call-taker recognition of stroke \parencite{cite6,cite7}. Call-takers who can rapidly and accurately recognise stroke are therefore crucial in facilitating prompt care in both pre-hospital and in-hospital settings.

@@

In this study, we use call recordings and registry data from the Copenhagen Emergency Medical Services (CEMS) and the Danish Stroke Registry (DanStroke) \parencite{cite16} from 2015 to 2020. We obtain call recordings from two call lines: the 1-1-2 emergency line and the medical helpline 1813 (MH-1813). We then fit a machine learning framework to classify medical helpline calls as stroke or non-stroke. Calls are first transcribed using an automatic speech recognition model and then categorised by a text classification model trained as an ensemble of five individual models. We compare the performance of the model with that of call-takers using MH-1813 data from 2021.

@@

We performed an occlusion analysis to evaluate the importance of individual words for both positive and negative classifier predictions \cref{tab_retrospective:table3-occlusion-analysis}. 
Among the words with a positive rank score, several words are synonymous with stroke, such as ``blood clot", ``haemorrhagic stroke", and ``stroke". Ambulances are rarely dispatched because the MH-1813 is not intended for emergencies. Therefore, a word like ``ambulance" may also be a strong indicator of call-taker recognition, which the model has learned to mimic. Additionally, most of the remaining words can be linked to stroke-related symptoms such as ``double vision", ``difficulties speaking", and ``hangs". Particularly, words describing the side of the body where symptoms occur ranked high (such as ``left", ``right", and ``side"). Finally, some words were also related to the sudden onset of symptoms (including ``suddenly" and ``minutes").

@@

\begin{table}[t]
    \centering
    \caption[Words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively.]{English translation of words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively. For this analysis, we used the model with the median F1-score out of 11 randomly seeded runs.}
    \label{tab_retrospective:table3-occlusion-analysis}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|lr|lr}
        \toprule
                    & \multicolumn{2}{c|}{Positive ranking score} & \multicolumn{2}{c}{Negative ranking score} \\
        \midrule
                    & \multicolumn{2}{c|}{Stroke predictions, $D=1,897$} & \multicolumn{2}{c}{Non-stroke predictions, $D=342,133$} \\
        \midrule
                    & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ \\
        \midrule
        1.  & Ambulance & 1,680 & Tetanus & 4,378 \\        
        2.  & Blood clot & 895 & Pregnant & 8,749 \\        
        3.  & Left & 1,108 & Cut & 7,592 \\        
        4.  & Right & 1,050 & Bandage & 4,561 \\        
        5.  & Double vision & 84 & Amager (a location) & 23,776 \\        
        6.  & The words & 344 & O'clock & 94,436 \\        
        7.  & Suddenly & 783 & The emergency room & 42,809 \\        
        8.  & Arm & 709 & The police & 2,903 \\        
        9.  & Side & 1,139 & Swollen & 60,559 \\        
        10. & Stroke & 117 & Over the counter (OTC) & 4,641 \\        
        11. & Double & 113 & The neck & 30,151 \\        
        12. & Control & 134 & Fever & 112,586 \\        
        13. & Call & 39 & Prescription & 5,450 \\        
        14. & Numb & 94 & Centimetre & 12,026 \\        
        15. & Minutes & 763 & The knee & 8,875 \\        
        16. & Difficulties speaking & 44 & The pharmacy & 10,085 \\        
        17. & Haemorrhagic stroke & 133 & The stomach & 42,105 \\        
        18. & Hand & 297 & Psychiatric & 3,688 \\        
        19. & The ambulance & 521 & Pneumonia & 7,597 \\        
        20. & Slurred speech & 58 & Stomach pain & 10,551 \\        
        21. & Blood clots & 224 & Stool & 19,155 \\        
        22. & Fast & 663 & The ribs & 3,928 \\        
        23. & Express & 44 & Bleed & 10,501 \\        
        24. & Blood thinner & 259 & Bleeding & 24,313 \\        
        25. & Incoherent & 15 & Ribs & 2,941 \\        
        26. & Lopsided & 211 & Broken & 19,415 \\        
        27. & Reduced & 528 & Inflammation & 10,050 \\        
        28. & Hangs & 628 & Common cold & 8,127 \\        
        29. & Transient & 48 & Morning or morrow & 78,558 \\        
        30. & Not making sense & 14 & Swelling & 17,762 \\        
        \bottomrule
    \end{tabular}%
    }
\end{table}

@@

This study does not imply that a machine learning model can replace medical call-takers. The effectiveness of the model is fully reliant on the conversation between the call-taker and caller and the call-taker's ability to skillfully triage the patient. Instead, the model should be used as a supportive tool for call-takers in the decision-making process, contributing to a higher recognition of patients with stroke and potentially boosting the confidence of call-takers in their decisions. A similar machine learning model designed to predict cardiac arrest was tested in a randomised controlled trial (RCT) at CEMS \parencite{cite15}. The results highlighted the necessity of incorporating input from call-takers. The machine learning model for cardiac arrest has subsequently been implemented in daily practice at CEMS, in a setup similar to the one presented in our study. However, implementation of our framework requires further investigation. The relative performance gap between call-takers and the model was larger in our study than in the cardiac arrest study \parencite{cite15}, which may affect the results of a potential RCT.

@@

The performance gap between the model and call-takers could be explained by the rarity of stroke calls to MH-1813 (0.250\% of all calls in 2021), which might affect call-taker awareness of stroke as a possible cause of certain symptoms. Additionally, certain stroke symptoms are so rare that some call-takers may never encounter them, increasing the risk of false negatives. The model was trained on more calls than any single call-taker would handle in a lifetime, enabling it to recognise even rare descriptors of stroke. The model is specifically trained to recognise strokes and exclusively learns from actual stroke descriptions, unlike call-takers, who are trained with generalised teaching materials to triage many different conditions. Therefore, call-takers may not have received specific training for patients with stroke and may never have encountered them.

@@

The improved sensitivity and PPV on the 65+ years group may be explained a higher prior probability of stroke for older patients and stronger evidence from the patient's medical history. The relatively high FOR and FPR for the 65+ group is likely to be a result of the much higher prevalence of stroke cases compared to the 18-64 year olds (0.85\% vs. 0.07\%). We did not have data to estimate potential bias related to race, ethnicity, language, accent, or dialects. Previous studies on speech recognition for call centres have indeed found that non-native speakers had a higher rate of transcription errors \parencite{han_deep_2017}. Since our model was trained on a representative - and therefore unbalanced - sample, we expect it to behave similarly. Future research should look to address these shortcomings, for example by utilising self"-supervised learning on massive amounts of diverse, unlabelled data covering multiple languages, accents, and dialects.

Due to European data regulations (GDPR), it was not possible to manually transcribe MH-1813 calls to train a new speech recognition model, so we had to rely on an existing solution. This also meant that we could not evaluate the word error rate (WER) of the model. Instead, we used the downstream performance of the text classification model when trained in combination with different speech recognition models to choose the best option. Since the focus of this study is the ability to correctly recognise stroke, and not the performance of the speech recognition model alone, this approach is better suited. Indeed, the WER might be misleading when choosing a speech recognition model for a specific task. For instance, one model might fail to predict redundant minimal response words (e.g. “uh” and “uhm”) and make small inflection errors (e.g. “clot” instead of “clots”), which results in a relatively high WER, while another model only fails to predict rare, specialised words that are highly indicative of stroke (e.g. “haemorrhage” and “thrombolysis”), which results in a relatively low WER.

Although we believe that the proposed machine learning framework can be further improved, several alternatives have already been explored in the preliminary experimental phase. The speech recognition model we used was trained on 1-1-2 calls for a previous project \parencite{cite14}, and so, was specialised to a domain very similar to that of MH-1813. We also tested an open-source, multilingual model from OpenAI called Whisper \parencite{radford_robust_2023}, but found that performance degraded slightly compared to the model trained on 1-1-2. We hypothesise that this is due to Whisper's inability to handle the specific noise conditions and recognise words from a specialised medical vocabulary.

For text classification, we used an ensemble of multi-layer perceptrons (MLPs). We also tested convolutional, recurrent, and self-attention (i.e. Transformer) architectures. However, this did not improve performance. In addition, we tested a pre"=trained self"=supervised model. Although many of these models are freely available to the public, they are primarily trained on English data. Only relatively few options exist for the Danish language, none of which are specialised in the medical domain. We used a monolingual Danish BERT model, which has previously been shown to outperform a multilingual alternative from Google for Danish named entity recognition \parencite{hvingelby_dane_2020}. However, this also did not result in a significant performance improvement. We hypothesise that the number of ground truth stroke positives was too small for these advanced models to learn more complex patterns than the MLP ensemble. In addition, a self"=supervised model would likely benefit from being pre"=trained on speech or text data from the target domain. Although training such large-scale foundation models have the potential to improve the classification model further, it is beyond the scope of this study. Thus, we chose the simpler MLP ensemble. We have included references to reviews of self"=supervised learning for speech and text in the references \parencite{gururangan_don_2020,mohamed_selfsupervised_2022}. Notably, it is not uncommon for small, simple models to match or outperform large, pre"=trained models for text-classification tasks \parencite{galke_bagofwords_2022}.

@@

Call-takers for both lines, who are nurses, paramedics, or physicians, can dispatch ambulances. The condition suspected by the call-taker is categorised based on a predefined diagnostic index and stored in an electronic record using a CAD system. The CAD records are associated with the Danish civil registration number (CPR number) \parencite{cite19} of the patient. The CPR number is a unique identification assigned to all Danish residents. It is used for interactions with health services and registries, enabling cross-referencing of the data sources used in this study. The call audio is recorded and stored separately from the CAD recordings using a telephone system.

@@

All patients with a final diagnosis of stroke or transient ischaemic attack admitted to a Danish hospital within 5 days of symptom onset are recorded in the Danish Stroke Registry \parencite{cite16}, also known as DanStroke. This record includes the patient-reported time of onset, stroke type (haemorrhagic, ischaemic, or transient ischaemic attack), and CPR number of the patient.
The diagnosis is obtained according to the national guidelines \parencite{baluenfeldt_national_2021}, which includes cerebral imaging and full diagnostic workup by neurologists. The validity of the Danish stroke registry has been shown to be high \parencite{wildenschild_registration_2013}, and the number of stroke mimics is therefore minimised in our dataset.

@@

The Danish Data Protection Agency (P-2021-475) approved this study. Danish law did not require approval from the Scientific Ethics Committee because the data were registry-based. CEMS approved the transcription of all calls made to 1-1-2 and MH-1813. All electronic records were anonymised before analysis, and the researchers did not inspect the calls manually.

@@

From the CAD medical records, we included all calls that could be matched to a corresponding audio file for 1-1-2 and MH-1813 from 2015 to 2021 for patients older than 18. The CAD records were matched with the telephone call recordings based on the call start, call duration, and call-taker identity. Due to data incompleteness, and the way the audio data is stored, at CEMS, 2,730,199 contacts could not be matched to their corresponding audiofile, however, 2,361,178 contacts were successfully matched. We found no obvious pattern in the matched and unmatched calls and we included all calls with at matching audio file. Next, a call was regarded as a case of ground truth stroke positive when the CPR number in the CAD record matched that of a DanStroke record, and the patient-reported time of onset was close to the call start time. We allowed a window of 72 hours before and 24 hours after the call starts to account for uncertainty in recording stroke onset time. We excluded calls involving subarachnoid haemorrhage cases. Finally, we considered a call to be a call-taker stroke positive when the call-taker selected the stroke diagnostic category during the call and dispatched an ambulance with the appropriate level of response \parencite{cite20}. To ensure that the effect of the machine learning framework was not overestimated, we excluded calls where diagnostic category had not been registered from the test set. We still reported the population characteristics and model performance of this group of calls to assess potential bias introduced by excluding them. A data-flow diagram is included in \cref{app:supplementary-retrospective}. The resulting dataset is the largest dataset of audio files from stroke calls collected to date.

@@

As the threshold differed for each model in the ensemble, computing the ensemble output score as the average output score of the individual models would not be meaningful. Instead, we first subtracted the threshold from the output score in the logit space (before sigmoid nonlinearity) for each model to obtain the same threshold (0.5). Subsequently, we defined the ensemble output score as the average of the centred output scores. The exact equations are provided in \cref{eq_retrospective:output_score,eq_retrospective:ensemble_prediction}.

@@

We performed an \emph{occlusion analysis} to better understand the predictions of the text classification model. This involved removing all instances of a given word from the input transcript to evaluate its impact on the model output. The word was removed before vectorisation, such that all word and character $n$-grams associated with the word were discarded. Specifically, let $z^{(n, d, w)}$ be the logit output of model $n$ in the ensemble for transcript $d$ when the word $w$ is occluded. For transcript $d$, we computed the word impact score $i^{(d, w)}$ as the mean difference between the logit before and after occlusion.
%
\begin{equation}
    i^{(d,w)} = \frac{1}{N_d} \sum_{n=1}^{N_d} \left( z^{(n, d)} - z^{(n, d, w)} \right) \enspace .
\end{equation}
%
We used the logit output to compute the impact score because the difference in sigmoid-normalised output is biased towards zero for values close to 0 or 1. To select words for inspection, we computed a ranking score, $r^{(w)}$, as the sum of the signed squares of the impact:
%
\begin{equation}
    r^{(w)} = \sum_{d=1}^{N} \text{sign}\left( i^{(d, w)} \right) \left( i^{(d,w)}\right) ^2 \enspace .
\end{equation}
%
where $\text{sign}$ represents the sign function. Squaring $i^{(d,w)}$ favours rare features with a high impact over common features with a low impact.

@@

We present the results with and without 1-1-2 training data, subgroup analyses based on age (18–64/65+) and sex (male/female), and call-takers performance. We also report the model performance on calls without a diagnostic category from the test year 2021 to assess potential data bias. We tested our results for statistical significance using approximate permutation tests. We used one-sided paired approximate permutation tests for model-to-model and model-to-call-taker comparisons when done on the same subset. For comparisons across different subsets (e.g. male vs. female) we used one-sided independent approximate permutation tests. We computed 95\% confidence intervals (CIs) using bootstrapping \parencite{dwass_modified_1957,eden_validity_1933}. In our assessment, we accounted for random variation associated with model training by basing the means, tests, and CIs on the predictions of 11 randomly initialised training runs. Statistical significance was defined as a p-value of less than 0.05.

@@

The source code can be shared using a Creative Commons NC-ND 4.0, an international licence, upon reasonable written request to the corresponding author, and requires a data use agreement.

@@

\section*{Declaration of interests}

JDH and LB received funding from the Innovation Fund Denmark. JDH and LB used Corti and hold stock warrants. LM is a co-founder, stockholder, and the Chief Technology Officer of Corti. JW received funding from Trygfonden. SNFB has no conflicts of interest to declare. HC has received funding from the Velux Foundation, Tværsfonden, Helsefonden, Hartmann Fonden, Lundbeck Foundation, and Novo Nordisk Foundation; royalties from Gyldendal; honoraria from Bayer and Bristol Meyers Squibb, and is chair of Action Plan for stroke in Europe Implementation, Co-chair of the Scientific Stroke Panel EAN and Senior Guest Editor of AHA Stroke. MS has no conflicts of interest to declare. HCC has no conflicts of interest to declare. CK received funding from the Novo Nordisk Foundation and is the chair of the Danish Resuscitation Council and vice chair of the Danish Stroke Society. Both positions are unpaid.

@@

JW, JDH and LB share lead authorship and contributed equally to the original drafting of the paper, review and editing, data curation, formal analysis, methodology, validation, and conceptualisation. JDH and LB contributed to software development and funding acquisition. SNFB contributed to the conceptualisation, data curation, funding acquisition, supervision, and writing - review and editing. LM contributed to conceptualisation, funding acquisition, project administration, supervision, and writing - review and editing. MS contributed to supervision and writing - review and editing. HC contributed to the methodology, supervision, and writing - review and editing. CK contributed to the conceptualisation, funding acquisition, methodology, project administration, supervision, and writing - review and editing. HCC contributed to the conceptualisation, data curation, funding acquisition, methodology, project administration, supervision, and writing - review and editing. HCC and CK share last authorship. The three lead authors directly accessed and verified the underlying data reported in this manuscript. JW's affiliations are exclusively academic.

>>>>

\textit{This chapter is a piece of original research published as part of the project:} \newline
\begin{center}
    \begin{enumerate}[leftmargin=8mm,rightmargin=8mm,topsep=0mm,label={[\Alph*]}]
        \setcounter{enumi}{5}
        \item \fullcite{wenstrup_retrospective_2023} \shared \hspace{0.1em} \parencite{wenstrup_retrospective_2023}
    \end{enumerate}
\end{center}

@@

\section*{Abstract}
Advanced stroke treatment is time dependent and, therefore, relies on recognition by call-takers at prehospital telehealth services to ensure fast hospitalization. This study aims to develop and assess the potential of machine learning in improving prehospital stroke recognition during medical helpline calls. We use calls from 1 January 2015 to 31 December 2020 in Copenhagen to develop a machine learning-based classification pipeline. Calls from 2021 are used for testing. Calls are first transcribed using an automatic speech recognition model and then categorized as stroke or non-stroke using a text classification model. Call-takers achieve a sensitivity of 52.7\% (95\% confidence interval 49.2-56.4\%) with a positive predictive value (PPV) of 17.1\% (15.5-18.6\%). The machine learning framework performs significantly better (p < 0.0001) with a sensitivity of 63.0\% (62.0-64.1\%) and a PPV of 24.9\% (24.3-25.5\%). Thus, a machine learning framework for recognizing stroke in prehospital medical helpline calls may become a supportive tool for call-takers, aiding in early and accurate stroke recognition.

\section{Introduction}
Stroke is a leading cause of disability and death worldwide \parencite{cite1,cite2,cite3}. Effective treatment is time-sensitive, and an optimal outcome is more likely when treatment is administered within the first four and a half hours from stroke onset \parencite{cite4,cite5}. The gateway to ambulance transport and hospital admittance is through prehospital telehealth services, including emergency medical call centers, nurse advice call lines, and out-of-hours health services. In the pre-hospital setting, the use of mobile stroke units has made it possible to deliver advanced treatment faster \parencite{cite6,cite7}. As the mobile stroke unit is only dispatched to patients with a suspected stroke, the impact of mobile stroke unit is directly influenced by accurate call-taker recognition of stroke \parencite{cite6,cite7}. Call-takers who can rapidly and accurately recognize stroke are therefore crucial in facilitating prompt care in both pre-hospital and in-hospital settings.

@@

In this study, we use call recordings and registry data from the Copenhagen Emergency Medical Services (CEMS) and the Danish Stroke Registry (DanStroke) \parencite{cite16} from 2015 to 2020. We obtain call recordings from two call lines: the 1-1-2 emergency line and the medical helpline 1813 (MH-1813). We then fit a machine learning framework to classify medical helpline calls as stroke or non-stroke. Calls are first transcribed using an automatic speech recognition model and then categorized by a text classification model trained as an ensemble of five individual models. We compare the performance of the model with that of call-takers using MH-1813 data from 2021.

@@

We performed an occlusion analysis to evaluate the importance of individual words for both positive and negative classifier predictions \cref{tab_retrospective:table3-occlusion-analysis}. 
Among the words with a positive rank score, several words are synonymous with stroke, such as ``blood clot", ``hemorrhagic stroke", and ``stroke". Ambulances are rarely dispatched because the MH-1813 is not intended for emergencies. Therefore, a word like ``ambulance" may also be a strong indicator of call-taker recognition, which the model has learned to mimic. Additionally, most of the remaining words can be linked to stroke-related symptoms such as ``double vision", ``difficulties speaking", and ``hangs". Particularly, words describing the side of the body where symptoms occur ranked high (such as ``left", ``right", and ``side"). Finally, some words were also related to the sudden onset of symptoms (including ``suddenly" and ``minutes").

@@

\begin{table}[t]
    \centering
    \caption[Words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively.]{English translation of words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively. For this analysis, we used the model with the median F1-score out of 11 randomly seeded runs.}
    \label{tab_retrospective:table3-occlusion-analysis}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|lr|lr}
        \toprule
                    & \multicolumn{2}{c|}{Positive ranking score} & \multicolumn{2}{c}{Negative ranking score} \\
        \midrule
                    & \multicolumn{2}{c|}{Stroke predictions, $D=1,897$} & \multicolumn{2}{c}{Non-stroke predictions, $D=342,133$} \\
        \midrule
                    & Word, $w$ \textit{(translated)} & Occurrences, $D^{(w)}$ & Word, $w$ \textit{(translated)} & Occurrences, $D^{(w)}$ \\
        \midrule
        1.  & Ambulance & 1,680 & Tetanus & 4,378 \\        
        2.  & Blood clot & 895 & Pregnant & 8,749 \\        
        3.  & Left & 1,108 & Cut & 7,592 \\        
        4.  & Right & 1,050 & Bandage & 4,561 \\        
        5.  & Double vision & 84 & Amager (a location) & 23,776 \\        
        6.  & The words & 344 & O'clock & 94,436 \\        
        7.  & Suddenly & 783 & The emergency room & 42,809 \\        
        8.  & Arm & 709 & The police & 2,903 \\        
        9.  & Side & 1,139 & Swollen & 60,559 \\        
        10. & Stroke & 117 & Over the counter (OTC) & 4,641 \\        
        11. & Double & 113 & The neck & 30,151 \\        
        12. & Control & 134 & Fever & 112,586 \\        
        13. & Call & 39 & Prescription & 5,450 \\        
        14. & Numb & 94 & Centimeter & 12,026 \\        
        15. & Minutes & 763 & The knee & 8,875 \\        
        16. & Difficulties speaking & 44 & The pharmacy & 10,085 \\        
        17. & Hemorrhagic stroke & 133 & The stomach & 42,105 \\        
        18. & Hand & 297 & Psychiatric & 3,688 \\        
        19. & The ambulance & 521 & Pneumonia & 7,597 \\        
        20. & Slurred speech & 58 & Stomach pain & 10,551 \\        
        21. & Blood clots & 224 & Stool & 19,155 \\        
        22. & Fast & 663 & The ribs & 3,928 \\        
        23. & Express & 44 & Bleed & 10,501 \\        
        24. & Blood thinner & 259 & Bleeding & 24,313 \\        
        25. & Incoherent & 15 & Ribs & 2,941 \\        
        26. & Lopsided & 211 & Broken & 19,415 \\        
        27. & Reduced & 528 & Inflammation & 10,050 \\        
        28. & Hangs & 628 & Common cold & 8,127 \\        
        29. & Transient & 48 & Morning or morrow & 78,558 \\        
        30. & Not making sense & 14 & Swelling & 17,762 \\        
        \bottomrule
    \end{tabular}%
    }
\end{table}

@@

This study does not imply that a machine learning model can replace medical call-takers. The effectiveness of the model is fully reliant on the conversation between the call-taker and caller and the call-taker's ability to skillfully triage the patient. Instead, the model should be used as a supportive tool for call-takers in the decision-making process, contributing to a higher recognition of patients with stroke and potentially boosting the confidence of call-takers in their decisions. A similar machine learning model designed to predict cardiac arrest was tested in a randomized controlled trial (RCT) at CEMS \parencite{cite15}. The results highlighted the necessity of incorporating input from call-takers. The machine learning model for cardiac arrest has subsequently been implemented in daily practice at CEMS, in a setup similar to the one presented in our study. However, implementation of our framework requires further investigation. The relative performance gap between call-takers and the model was larger in our study than in the cardiac arrest study \parencite{cite15}, which may affect the results of a potential RCT.

@@

The performance gap between the model and call-takers could be explained by the rarity of stroke calls to MH-1813 (0.250\% of all calls in 2021), which might affect call-taker awareness of stroke as a possible cause of certain symptoms. Additionally, certain stroke symptoms are so rare that some call-takers may never encounter them, increasing the risk of false negatives. The model was trained on more calls than any single call-taker would handle in a lifetime, enabling it to recognize even rare descriptors of stroke. The model is specifically trained to recognize strokes and exclusively learns from actual stroke descriptions, unlike call-takers, who are trained with generalized teaching materials to triage many different conditions. Therefore, call-takers may not have received specific training for patients with stroke and may never have encountered them.

@@

The improved sensitivity and PPV on the 65+ years group may be explained a higher prior probability of stroke for older patients and stronger evidence from the patient's medical history. The relatively high FOR and FPR for the 65+ group is likely to be a result of the much higher prevalence of stroke cases compared to the 18-64 year olds (0.85\% vs. 0.07\%). We did not have data to estimate potential bias related to race, ethnicity, language, accent, or dialects. Previous studies on speech recognition for call centers have indeed found that non-native speakers had a higher rate of transcription errors \parencite{han_deep_2017}. Since our model was trained on a representative - and therefore unbalanced - sample, we expect it to behave similarly. Future research should look to address these shortcomings, for example by utilizing self"-supervised learning on massive amounts of diverse, unlabeled data covering multiple languages, accents, and dialects.

Due to European data regulations (GDPR), it was not possible to manually transcribe MH-1813 calls to train a new speech recognition model, so we had to rely on an existing solution. This also meant that we could not evaluate the word error rate (WER) of the model. Instead, we used the downstream performance of the text classification model when trained in combination with different speech recognition models to choose the best option. Since the focus of this study is the ability to correctly recognize stroke, and not the performance of the speech recognition model alone, this approach is better suited. Indeed, the WER might be misleading when choosing a speech recognition model for a specific task. For instance, one model might fail to predict redundant minimal response words (e.g. “uh” and “uhm”) and make small inflection errors (e.g. “clot” instead of “clots”), which results in a relatively high WER, while another model only fails to predict rare, specialized words that are highly indicative of stroke (e.g. “hemorrhage” and “thrombolysis”), which results in a relatively low WER.

Although we believe that the proposed machine learning framework can be further improved, several alternatives have already been explored in the preliminary experimental phase. The speech recognition model we used was trained on 1-1-2 calls for a previous project \parencite{cite14}, and so, was specialized to a domain very similar to that of MH-1813. We also tested an open-source, multilingual model from OpenAI called Whisper \parencite{radford_robust_2023}, but found that performance degraded slightly compared to the model trained on 1-1-2. We hypothesize that this is due to Whisper's inability to handle the specific noise conditions and recognize words from a specialized medical vocabulary.

For text classification, we used an ensemble of multi-layer perceptrons (MLPs). We also tested convolutional, recurrent, and self-attention (i.e. Transformer) architectures. However, this did not improve performance. In addition, we tested a pre"=trained self"=supervised model. Although many of these models are freely available to the public, they are primarily trained on English data. Only relatively few options exist for the Danish language, none of which are specialized in the medical domain. We used a monolingual Danish BERT model, which has previously been shown to outperform a multilingual alternative from Google for Danish named entity recognition \parencite{hvingelby_dane_2020}. However, this also did not result in a significant performance improvement. We hypothesize that the number of ground truth stroke positives was too small for these advanced models to learn more complex patterns than the MLP ensemble. In addition, a self"=supervised model would likely benefit from being pre"=trained on speech or text data from the target domain. Although training such large-scale foundation models have the potential to improve the classification model further, it is beyond the scope of this study. Thus, we chose the simpler MLP ensemble. We have included references to reviews of self"=supervised learning for speech and text in the references \parencite{gururangan_don_2020,mohamed_selfsupervised_2022}. Notably, it is not uncommon for small, simple models to match or outperform large, pre"=trained models for text-classification tasks \parencite{galke_bagofwords_2022}.

@@

Call-takers for both lines, who are nurses, paramedics, or physicians, can dispatch ambulances. The condition suspected by the call-taker is categorized based on a predefined diagnostic index and stored in an electronic record using a CAD system. The CAD records are associated with the Danish civil registration number (CPR number) \parencite{cite19} of the patient. The CPR number is a unique identification assigned to all Danish residents. It is used for interactions with health services and registries, enabling cross-referencing of the data sources used in this study. The call audio is recorded and stored separately from the CAD recordings using a telephone system.

@@

All patients with a final diagnosis of stroke or transient ischemic attack admitted to a Danish hospital within 5 days of symptom onset are recorded in the Danish Stroke Registry \parencite{cite16}, also known as DanStroke. This record includes the patient-reported time of onset, stroke type (hemorrhagic, ischemic, or transient ischemic attack), and CPR number of the patient.
The diagnosis is obtained according to the national guidelines \parencite{baluenfeldt_national_2021}, which includes cerebral imaging and full diagnostic workup by neurologists. The validity of the Danish stroke registry has been shown to be high \parencite{wildenschild_registration_2013}, and the number of stroke mimics is therefore minimized in our dataset.

@@

The Danish Data Protection Agency (P-2021-475) approved this study. Danish law did not require approval from the Scientific Ethics Committee because the data were registry-based. CEMS approved the transcription of all calls made to 1-1-2 and MH-1813. All electronic records were anonymized before analysis, and the researchers did not inspect the calls manually.

@@

From the CAD medical records, we included all calls that could be matched to a corresponding audio file for 1-1-2 and MH-1813 from 2015 to 2021 for patients older than 18. The CAD records were matched with the telephone call recordings based on the call start, call duration, and call-taker identity. Due to data incompleteness, and the way the audio data is stored, at CEMS, 2,730,199 contacts could not be matched to their corresponding audio file, however, 2,361,178 contacts were successfully matched. We found no obvious pattern in the matched and unmatched calls and we included all calls with at matching audio file. Next, a call was regarded as a case of ground truth stroke positive when the CPR number in the CAD record matched that of a DanStroke record, and the patient-reported time of onset was close to the call start time. We allowed a window of 72 hours before and 24 hours after the call starts to account for uncertainty in recording stroke onset time. We excluded calls involving subarachnoid hemorrhage cases. Finally, we considered a call to be a call-taker stroke positive when the call-taker selected the stroke diagnostic category during the call and dispatched an ambulance with the appropriate level of response \parencite{cite20}. To ensure that the effect of the machine learning framework was not overestimated, we excluded calls where diagnostic category had not been registered from the test set. We still reported the population characteristics and model performance of this group of calls to assess potential bias introduced by excluding them. A data-flow diagram is included in \cref{app:supplementary-retrospective}. The resulting dataset is the largest dataset of audio files from stroke calls collected to date.

@@

As the threshold differed for each model in the ensemble, computing the ensemble output score as the average output score of the individual models would not be meaningful. Instead, we first subtracted the threshold from the output score in the logit space (before sigmoid nonlinearity) for each model to obtain the same threshold (0.5). Subsequently, we defined the ensemble output score as the average of the centered output scores. The exact equations are provided in \cref{eq_retrospective:output_score,eq_retrospective:ensemble_prediction}.

@@

We performed an \emph{occlusion analysis} to better understand the predictions of the text classification model. This involved removing all instances of a given word from the input transcript to evaluate its impact on the model output. The word was removed before vectorization, such that all word and character $n$-grams associated with the word were discarded. Specifically, let $z^{(n, d, w)}$ be the logit output of model $n$ in the ensemble for transcript $d$ when the word $w$ is occluded. For transcript $d$, we computed the word impact score $i^{(d, w)}$ as the mean difference between the logit before and after occlusion.
%
\begin{equation}
    i^{(d,w)} = \frac{1}{N_d} \sum_{n=1}^{N_d} \left( z^{(n, d)} - z^{(n, d, w)} \right) \enspace .
\end{equation}
%
We used the logit output to compute the impact score because the difference in sigmoid-normalised output is biased towards zero for values close to 0 or 1. To select words for inspection, we computed a ranking score, $r^{(w)}$, as the sum of the signed squares of the impact:
%
\begin{equation}
    r^{(w)} = \sum_{d=1}^{N} \text{sign}\left( i^{(d, w)} \right) \left( i^{(d,w)}\right) ^2 \enspace .
\end{equation}
%
where $\text{sign}$ represents the sign function. Squaring $i^{(d,w)}$ favors rare features with a high impact over common features with a low impact.

@@

We present the results with and without 1-1-2 training data, subgroup analyses based on age (18–64/65+) and sex (male/female), and call-takers performance. We also report the model performance on calls without a diagnostic category from the test year 2021 to assess potential data bias. We tested our results for statistical significance using approximate permutation tests. We used one-sided paired approximate permutation tests for model-to-model and model-to-call-taker comparisons when done on the same subset. For comparisons across different subsets (e.g. male vs. female) we used one-sided independent approximate permutation tests. We computed 95\% confidence intervals (CIs) using bootstrapping \parencite{dwass_modified_1957,eden_validity_1933}. In our assessment, we accounted for random variation associated with model training by basing the means, tests, and CIs on the predictions of 11 randomly initialized training runs. Statistical significance was defined as a p-value of less than 0.05.

@@

The source code can be shared using a Creative Commons NC-ND 4.0, an international license, upon reasonable written request to the corresponding author, and requires a data use agreement.

@@

@@

JW, JDH and LB share lead authorship and contributed equally to the original drafting of the paper, review and editing, data curation, formal analysis, methodology, validation, and conceptualization. JDH and LB contributed to software development and funding acquisition. SNFB contributed to the conceptualization, data curation, funding acquisition, supervision, and writing - review and editing. LM contributed to conceptualization, funding acquisition, project administration, supervision, and writing - review and editing. MS contributed to supervision and writing - review and editing. HC contributed to the methodology, supervision, and writing - review and editing. CK contributed to the conceptualization, funding acquisition, methodology, project administration, supervision, and writing - review and editing. HCC contributed to the conceptualization, data curation, funding acquisition, methodology, project administration, supervision, and writing - review and editing. HCC and CK share last authorship. The three lead authors directly accessed and verified the underlying data reported in this manuscript. JW's affiliations are exclusively academic.
